The Proofs

### 2023-10-24 proofs/main.tex
WIP: Algorithm

Let us overview the inference rules of the subtyping algorithm.
\begin{itemize}
  \item \ruleref{\ottdruleANVarLabel} and \ruleref{\ottdruleAPVarLabel} 
    are the base cases. They copy the corresponding declarative rules and
    ensure the reflexivity.
  \item \ruleref{\ottdruleAPUVarLabel} is the only case generating 
    subtyping constraints. In this case, we must ensure
    that the resulting constrains guarantees that the instantiation of 
    $[[â⁺]]$ is a supertype of $[[iP]]$.
    However, the obvious constraint $[[â⁺ :≥ iP]]$ might be problematic
    if  $[[iP]]$ is not well-formed in $[[Θ(â⁺)]]$. For this reason,
    we use the \emph{upgrade} procedure (it will be covered in \cref{sec:lub})
    to find the minimal supertype of $[[iP]]$, which is well-formed in $[[Θ(â⁺)]]$.

    Notice that this rule does not have a negative counterpart. This is
    because one one of the important invariants of the algorithm: 
    in the negative subtyping, only positive algorithmic variables
    can occur in the types.

  \item \ruleref{\ottdruleAShiftDLabel} and \ruleref{\ottdruleAShiftULabel} are the
    \emph{shift} rules. According to the declarative system,
    shifted subtyping requires equivalence. In presence of the algorithmic 
    variables, it means that the left and the right-hand sides of the subtyping
    must be unified. Hence, the shift rules invoke the unification algorithm, 
    which will be discussed in \cref{sec:unification}. The unification 
    returns the minimal set of constraints $[[UC]]$, which is necessary
    and sufficient for the subtyping.

  \item \ruleref{\ottdruleAArrowLabel}.
    In this case, the algorithm makes two calls:
    a recursive call to the negative subtyping algorithm for the argument types,
    and a call to the positive subtyping algorithm for the result types.
    After that, the resulting constraints are merged using the
    \emph{subtyping constraint merge} procedure, 
    which is discussed in \cref{sec:constraint-merge}.
  \item \ruleref{\ottdruleAForallLabel} and \ruleref{\ottdruleAExistsLabel}
    are symmetric. These are the only places where 
    the algorithmic variables are introduced.
    It is done by algorithmization (\cref{sec:variable-algorithmization}) 
    of the quantified variables: these variables are replaced by 
    fresh algorithmic variables in the body of the quantified type,
    the algorithmic variables are added to the subtyping context $[[Θ]]$,
    after that the recursive call is made. Notice that the declarative context
    $[[Γ]]$ is extended by the quantified variables from the right-hand side,
    which matches the declarative system.
\end{itemize}

@@

Unification and subtyping algorithms are based on the constraint generation.
The constraints are represented by set of constraint entries.

@@

\begin{algorithm}[Merge of Subtyping Constraints]
  \label{definition:merge-subtyping-constraints}
  Suppose that $[[Θ ⊢ SC1]]$ and $[[Θ ⊢ SC2]]$.
  Then $[[Θ ⊢ SC1 & SC2 = SC]]$
  defines a set such that $[[ucE]] \in [[SC]]$ iff either
  \begin{itemize}
    \item $[[ucE]] \in [[SC1]]$ and there is no matching $[[ucE']] \in [[SC2]]$; or
    \item $[[ucE]] \in [[SC2]]$ and there is no matching $[[ucE']] \in [[SC1]]$; or
    \item $[[Θ(α̂±) ⊢ ucE1 & ucE2 = ucE]]$ for some $[[ucE1]] \in [[SC1]]$ and $[[ucE2]] \in [[SC2]]$
      such that $[[ucE1]]$ matches with $[[ucE2]]$ restricting variable
      $[[α̂±]]$. 
  \end{itemize}
\end{algorithm}

>>>>

Let us overview the inference rules of the subtyping algorithm.
\begin{itemize}
  \item \ruleref{\ottdruleANVarLabel} and \ruleref{\ottdruleAPVarLabel} 
    are the base cases. They copy the corresponding declarative rules and
    ensure reflexivity.
  \item \ruleref{\ottdruleAPUVarLabel} is the only case generating 
    subtyping constraints. In this case, we must ensure
    that the resulting constraints guarantee that the instantiation of 
    $[[â⁺]]$ is a supertype of $[[iP]]$.
    However, the obvious constraint $[[â⁺ :≥ iP]]$ might be problematic
    if  $[[iP]]$ is not well-formed in $[[Θ(â⁺)]]$. For this reason,
    we use the \emph{upgrade} procedure (it will be covered in \cref{sec:lub})
    to find the minimal supertype of $[[iP]]$, which is well-formed in $[[Θ(â⁺)]]$.

    Notice that this rule does not have a negative counterpart. This is
    because one of the important invariants of the algorithm: 
    in the negative subtyping, only positive algorithmic variables
    can occur in the types.

  \item \ruleref{\ottdruleAShiftDLabel} and \ruleref{\ottdruleAShiftULabel} are the
    \emph{shift} rules. According to the declarative system,
    shifted subtyping requires equivalence. In the presence of the algorithmic 
    variables, it means that the left and the right-hand sides of the subtyping
    must be unified. Hence, the shift rules invoke the unification algorithm, 
    which will be discussed in \cref{sec:unification}. The unification 
    returns the minimal set of constraints $[[UC]]$, which is necessary
    and sufficient for the subtyping.

  \item \ruleref{\ottdruleAArrowLabel}.
    In this case, the algorithm makes two calls:
    a recursive call to the negative subtyping algorithm for the argument types,
    and a call to the positive subtyping algorithm for the result types.
    After that, the resulting constraints are merged using the
    \emph{subtyping constraint merge} procedure, 
    which is discussed in \cref{sec:constraint-merge}.
  \item \ruleref{\ottdruleAForallLabel} and \ruleref{\ottdruleAExistsLabel}
    are symmetric. These are the only places where 
    the algorithmic variables are introduced.
    It is done by algorithmization (\cref{sec:variable-algorithmization}) 
    of the quantified variables: these variables are replaced by 
    fresh algorithmic variables in the body of the quantified type,
    the algorithmic variables are added to the subtyping context $[[Θ]]$,
    after that, the recursive call is made. Notice that the declarative context
    $[[Γ]]$ is extended by the quantified variables from the right-hand side,
    which matches the declarative system.
\end{itemize}

@@

Unification and subtyping algorithms are based on constraint generation.
The constraints are represented by a set of constraint entries.

@@

\begin{algorithm}[Merge of Subtyping Constraints]
  \label{definition:merge-subtyping-constraints}
  Suppose that $[[Θ ⊢ SC1]]$ and $[[Θ ⊢ SC2]]$.\\
  Then $[[Θ ⊢ SC1 & SC2 = SC]]$
  defines a set of constraints $[[SC]]$ such that $[[scE]] \in [[SC]]$ iff either:
  \begin{itemize}
    \item $[[scE]] \in [[SC1]]$ and there is no matching $[[scE']] \in [[SC2]]$; or
    \item $[[scE]] \in [[SC2]]$ and there is no matching $[[scE']] \in [[SC1]]$; or
    \item $[[Θ(α̂±) ⊢ scE1 & scE2 = scE]]$ for some $[[scE1]] \in [[SC1]]$ and $[[scE2]] \in [[SC2]]$
      such that $[[scE1]]$ and $[[scE2]]$ both restrict variable $[[α̂±]]$. 
  \end{itemize}
\end{algorithm}
